---
title: "Calculating community measures"
author: "Jarad Niemi"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Calculate abundance, richness, and diversity from count data. 

## Combine survey and count data

In the situation that some surveys had no birds observed, there will be no
entry in the `avian_common_count` data set. 
Thus, we combine this data set with the `avian_surveys` data set to insure that
all surveys are represented. 
The following step, which fills in zeros for any species not observed during a
particular survey, is unnecessary for calculation of abundance, richness, and
diversity but will likely be helpful for other tasks, e.g. abundance or 
occupancy modeling. 

```{r load_data}
library(dplyr)
library(tidyr)
library(STRIPSSchulte)
all_counts <- avian_common_count %>%
  right_join(avian_surveys, by="survey") %>%     # include all surveys
  complete(survey, abbrev, fill = list(count=0)) # fill with zero counts
```



## Calculate abundance, richness, and diversity

The [vegan](https://cran.r-project.org/web/packages/vegan/index.html) package
has functionality to calculate Shannon, Simpson, and inverse Simpson measures
of diversity. 
For avian populations, we typically use inverse Simpson diversity, but there is
an issue with defining diversity when no individuals of any species are 
observed. 
We propose (perhaps it is already used or there is a different solution) to 
define inverse Simpson diversity to be 0 when there are no individuals of any 
species observed. 

```{r}
diversity <- function(x) {
  div <- vegan::diversity(x, index = "inv")
  div[is.infinite(div)] <- 0
  return(div)
}
```

```{r}
all_counts %>%
  group_by(survey) %>%
  summarize(abundance = sum(count),
            richness  = sum(count>0),
            diversity = diversity(count)) # Inverse Simpson's Diversity
```

